---
title: "2: Formatting and Visualising BridgeAI EoI forms"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,       # Show the code in the html
  results = "hide" # Hide the results in the html
)
```

# Step 1 - Define location of input files

This notebook will process the current EoI batch, outputting a formatted excel file. It assumes you have all your EoI inputs inside a folder called 'Inputs'. During the processing it will compare the current EoI to all previous EoIs within the same Input folder, to check for duplicates.

EDIT BELOW TO MATCH FILES ON YOUR COMPUTER:

```{r}
# directory (folder) that contains your 'Inputs' folder:
cEoI_directory <- "~/Documents/GitHub/BridgeAI/EoIs"
# EoI file you want to process:
cEoI_csv_filename <- "EoI_Demo_B1.csv"
```

Now run this code chunk below to check these inputs are valid:

```{r}
source("sub_functions.R") #  source the sub-functions, needed for rest of this notebook
check_inputs(cEoI_directory,cEoI_csv_filename)
```

# Step 2 - Read EoI input & create output directories

## 2a. Load in the current EoI (cEoI)

Running the code chunk below will load in the current EoI (the one you want to process)

```{r}
cEoI_file_path <- paste0(cEoI_directory, '/Inputs/', cEoI_csv_filename) 
cEoI <- read.csv(cEoI_file_path, header = TRUE) # Read the current EoI CSV file into R
raw_cEoI <- cEoI # to save for later, unedited
head(cEoI) # view the first few rows of the data
```

If you want to have a broader view of the data, go to your 'Environment tab' in RStudio and click on 'cEoI' - this will open up a new window for you to browse. Double check that the data has been read into R correctly i.e. correct size, not corrupted (numbers and dates read in as expected).

## 2b. Create output directories (to save output files)

```{r}

parent_output_dir <- paste0(cEoI_directory, '/Outputs') 

# Create parent_output_dir if it does not exist yet
if (!dir.exists(parent_output_dir)) { 
  dir.create(parent_output_dir)
}

# Define the specific output directory for this EoI session
output_dir <- paste0(parent_output_dir, '/OUTPUTS_', tools::file_path_sans_ext(basename(cEoI_csv_filename)))

# Create output_dir if it does not exist yet
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
} else {
  warning(paste0('Output directory already exists for this EoI(', cEoI_csv_filename, '). Running the rest of the code may mean previously created files get overwritten.'))
}
```

# Step 3 - Validation of cEoI

## 3a. Check that the csv structure matches the demo structure

We are assuming that the folder 'Demo_Inputs' is in the same location as this 1_EoI_Process.Rmd file, and contains the file 'EoI_Demo_B1.csv'.

```{r}
# run function to check for demo match
demo_match(cEoI_df = cEoI)
```

## 3b. No Consent

Find if any companies did not consent to their data being processed. If companies are found to have selected 'No', they will be put into a new dataframe 'cEoI_NoConsent', and the rest of the companies are put into 'cEoI_Val1', indicating they have passed validation step 1.

```{r}
# run a function that finds if any company did not consent
val_1_results <- val_1(cEoI_df = cEoI)
# view the results - NoConsent
head(val_1_results$cEoI_NoConsent)
# view the results - Validation 1
```

## 3c. Duplicates

Check if there are any duplicate companies across batches . The code chunk below compares the currently select EoI with any other EoIs it finds in the 'Inputs' directory. If duplicates are found they will be put into a new dataframe 'cEoI_Duplicates', and the rest of the companies are put into 'cEoI_Val12', indicating they have passed validation step 1 and 2.

```{r}
# run a function that spots duplicates
val_2_results <- val_2(cEoI_directory,cEoI_csv_filename,val_1_results$cEoI_Val1)

```

## 3d. Validate the validations!

We now have 3 dataframes (cEoI_Val12, cEoI_NoConsent and cEoI_Duplicates). The number of rows (companies) in these dataframes should add up to the same number in the original dataframe (cEoI) before we did anything to it. Let's check that, to make sure we have not lost or added a company.

```{r}
# run a function to validate the validations! 
val_check(val_2_results$cEoI_Val12,val_1_results$cEoI_NoConsent,val_2_results$cEoI_Duplicates)
```

Take a look at the validated data frame:

```{r}
head(val_2_results$cEoI_Val12)
```

# Step 4 - Tidy up Current_EoI

Running the code chunk below will rename columns to be more sensible to read.

-   Step a. Replace '.' with a space (works better for some row names than others)

-   Step b. Rename some specific columns (feel free to change this!)

-   Step c. Add in new columns, that will be filled in later (feel free to change this!)

-   Step d. Remove some columns that will be less relevant when pairing with ISAs (feel free to change!)

```{r}
# run a function to tidy up current EoI
tidy_cEoI <- tidy_up(cEoI_df, val_2_results$cEoI_Val12, val_2_results$cEoI_Duplicates, val_1_results$cEoI_NoConsent)
```

Explore the result of each tidy data frame, for example:

```{r}
head(tidy_cEoI$cEoI_Val12)
```

# Step 5 - Restructure validated cEoI and save to excel

Now we will restructure the dataframe(s) in a way that is helpful for the next steps of matching each company to an ISA. We will then save each dataframe as a separate tab of an excel output.

```{r}
# run a function that restructures and saves to excel
save_to_excel(raw_cEoI,tidy_cEoI$cEoI_Val12,tidy_cEoI$cEoI_NoConsent,tidy_cEoI$cEoI_Duplicate,cEoI_csv_filename,output_dir)
```
